---
title: "Mapping & Mining Cash Based Intervention Process"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      collapse = FALSE,
                      comment = "#>",
                      fig.align = "center")
knitr::opts_chunk$set(fig.width = 9, 
                      fig.height = 7)
set.seed(1)
extrafont::loadfonts(quiet=TRUE)
options(scipen = 999) # turn-off scientific notation like 1e+48
library(tidyverse)
library(bupaR)
library(edeaR)
library(eventdataR)
library(processmapR)
library(processmonitR)
library(xesreadR)
library(petrinetR)
#devtools::install_github("process-analytics/bpmn-visualization-R")
library(bpmnVisualization) 
```

> “It is not enough to do your best; you must know what to do, and then do your best.”
> “If you can't describe what you are doing as a process, you don't know what you're doin" -- Edward Deming


    
## What & Why Mapping process? 
 *  is a hierarchical method for displaying processes that illustrates how a product or transaction is processed.​

 *  is a visual representation of the work-flow either within a process - or an image of the whole operation.​

 *  should allow people unfamiliar with the process to understand the interaction of causes during the work-flow.

Why ?

 * Efficiency    
 * Information gap 
 
 * Bottlenecks  
 
 * Length & delays  
 
 * Ownership 
 
 * Roles and responsibility  
 * Risks management  
 * Trends & Comparison


## Process Mining 

Mining is much more than using a specific tool. Mostly, it is an iterative procedure involving asking the relevant business questions, understanding the data, interpreting the data correctly (statistical significance vs.  practical relevance), and most importantly deriving measures for improving the process under investigation.

```{r}
bpmn_file <- here::here("data-raw", "CBI.bpmn")
 
bpmnVisualization::display(bpmn_file)
``` 

### Intervention Events

I order to fit for process mining, we need __event log__. The analysis need a specific class of object (build through the function bupaR::eventlog) whcih requires the following:  
 

 * __case_id__ : The case to which the event belongs. (A character vector containing variable names of length 1 or more).

 * __activity_id__ : The activity the event refers to. (A character vector containing variable names of length 1 or more).
 
   * Outreached
   * Intake Registration
   * Post-Distribution Monitored
   * Assessed as Eligible
   * Assessed as Not Eligible
   * Appealed
   * Re-assessed
   * Enrolled / prioritised
   * Not-enrolled
   * Notified
   * Assisted
   

 * __activity_instance_id__ : The stage in the transactional life cycce.

 * __lifecycle_id__ : The life cycle classifier of the event log.

 * __timestamp__ : The timestamp of the event. (Should refer to a Date or POSIXct field).

 * __resource_id__ : The resource identifier of the event log. This can be a staff - a budget or a contract service for instances. A character vector containing variable names of length 1 or more.

 * __order__ : Configure how to handle sort events with equal timestamps: auto will use the order in the original data, alphabetical will sort the activity labels by alphabet, sorted will assume that the data frame is already correctly sorted and has a column '.order', providing a column name will use this column for ordering (can be numeric of character). The latter will never overrule timestamp orderings.

	
When validate is 'TRUE' some basic checks are run on the contents of the event log such as that activity instances are not connected to more than one case or activity. Using 'FALSE' improves the performance by skipping those checks.

```{r}
data("patients")
#class(patients)
#head(patients)
pat <-  patients

pat$process_step <- recode(pat$handling,         
                            "Registration"="IntakeRegistration / Self-Targeting",         
                            "Triage and Assessment"="vulnerability Assessment",
                            "Blood test" = "Assessed as Eligible",       
                            "MRI SCAN"  = "Notified" , 
                            "X-Ray" = "Appealed",          
                            "Discuss Results" = "Received Assistnce" ,
                            "Check-out" = "Post Distribution Monitoring"           )
#levels(as.factor(pat$handling))
pat$lifecycle <- ifelse(pat$handling == "check-out", "Complete", "Incomplete")
pat$activity_instance_id <- row.names(pat)

myeventlog <- bupaR::eventlog(pat,
                case_id = "patient",
                activity_id = "process_step",
                activity_instance_id = "handling_id",
              #  lifecycle_id = "lifecycle",
                timestamp = "time",
                resource_id = "employee",
                 validate = TRUE)
```

 
### Assessing Process Flow Level 
```{r}
processmapR::process_map(myeventlog)
```

```{r}
myeventlog %>%
  filter_activity_frequency(percentage = 1.0) %>% 
  filter_trace_frequency(percentage = .30) %>%    
  processmapR::process_map(performance(mean, "mins"), render = T)
```

### What are the Activites Sequence?

```{r}
precedence_matrix <- myeventlog  %>%
  filter_activity_frequency(percentage = 1.0) %>% 
  filter_trace_frequency(percentage = .80) %>%    
  processmapR::precedence_matrix() %>% 
  plot()
precedence_matrix
```



```{r}
# ## Show throughput time; In hours by Application Type  
# myeventlog %>%
#   # show only the most frequent traces
#   edeaR::filter_trace_frequency(percentage = .80) %>%    
#   group_by(registration_type) %>% 
#   # Metric: Throughput time of cases
#   edeaR::throughput_time('log', units = 'hours')
```


### What are the Frequence of Activities?

```{r}
myeventlog %>%
				activity_frequency("activity") %>%
				plot() 
```
 				   		 
			 
```{r}
### Activity Presence
# myeventlog %>%
# 				activity_presence() %>%
# 				plot() 

```

### How much resources have been used?		 

```{r} 
plot(edeaR::end_activities(myeventlog, level_of_analysis="resource"))
```

### Resource Map 

```{r}
processmapR::resource_map(myeventlog)
```


### Resource frequency 

```{r}
myeventlog %>%
					resource_frequency("log") %>%
					plot()
```
 

```{r}
myeventlog %>%
					resource_frequency("case") %>%
					plot()
```
 

```{r}
myeventlog %>%
					resource_frequency("resource") %>%
					plot()
```
 

```{r}
myeventlog %>%
					resource_frequency("activity") %>%
					plot()
```
 

```{r}
myeventlog %>%
					resource_frequency("resource-activity") %>%
					plot()
```

### Resource involvement 

```{r}
myeventlog %>%
					resource_involvement("resource") %>%
					plot()
```
 

```{r}
myeventlog %>%
					resource_involvement("case") %>%
					plot()
```
 

```{r}
myeventlog %>%
					resource_involvement("resource-activity") %>%
					plot()
```

### Resource specialization 

```{r}
myeventlog %>%
					resource_specialisation("log") %>%
					plot()
```
 
```{r}
myeventlog %>%
					resource_specialisation("resource") %>%
					plot()
```
 

```{r}
myeventlog %>%
					resource_specialisation("activity") %>%
					plot()
```


## Performance

### Relative Start in Hours

```{r}
processmapR::dotted_chart(myeventlog,
                          x = "relative", 
                          y ="start",
                          color = NULL, units ="hours")
```


### Throughput Time 
```{r}

#("units", "Time units:", choices = c("min","hours","days","weeks"), selected = "hours")
#input$units <- "hours"

myeventlog %>%
				throughput_time("log", units = "hours") %>%
				plot() 
```
 				   		  

```{r}
myeventlog %>%
				throughput_time("case", units = "hours") %>%
				plot() 
```
 				   		 
			 

### Processing Time 

```{r}
myeventlog %>%
				processing_time("log", units = "hours") %>%
				plot() 
```
 				   		 

```{r}
myeventlog %>%
				processing_time("case", units = "hours") %>%
				plot() 
```
 				   		 
			  

```{r}
myeventlog %>%
				processing_time("activity", units = "hours") %>%
				plot() 
```
 				   		  

```{r}
myeventlog %>%
				processing_time("resource", units = "hours") %>%
				plot() 
```
 				   		 
			 
### Idle Time 

```{r}
myeventlog %>%
				idle_time("log", units = "hours") %>%
				plot() 
```
 				   		  

```{r}
myeventlog %>%
				idle_time("case", units = "hours") %>%
				plot() 
```
 				   		 
```{r}
myeventlog %>%
				idle_time("resource", units = "hours") %>%
				plot() 
```
 				   		 
			  

## Rework

## Selfloops Matrix

```{r}
myeventlog %>%
				redo_selfloops_referral_matrix() %>%
				plot()
```


## Repetitions Matrix
```{r}
myeventlog %>%
				redo_repetitions_referral_matrix() %>%
				plot()

		
```  
  
