---
title: "BPMN Visualization - R Package"
output: html_document
--- 
This project is an R package for visualizing process execution data on BPMN diagrams, using overlays, style customization and interactions.

It is made possible by [htmlwidgets](http://www.htmlwidgets.org/), which provides an easy-to-use framework for bringing together R and the [BPMN Visualization](https://github.com/process-analytics/bpmn-visualization-js) TypeScript library.
 
 
 https://process-analytics.github.io/bpmn-visualization-js/#_live_examples
 
https://medium.com/@process-analytics/hacktoberfest-2021-with-process-analytics-44eecc238ead 
Build your process map (i.e a .bpmn file ) with Camunda - https://camunda.com/download/modeler/ 

#### Install _BPMN Visualization - R Package_ from GitHub

 
```{r}
#devtools::install_github("process-analytics/bpmn-visualization-R")
```

Then, make the library available to the current R project:
```{r}
library(bpmnVisualization)
```

### Load the BPMN file
`bpmnVisualization` accepts a BPMN file (or an XML document):

  The package provides some [BPMN examples](./inst/examples). \
They are taken from the [BPMN 2.0 examples non-normative machine readable files](https://www.omg.org/spec/BPMN/2.0).
- Email_Voting.bpmn
- Nobel_Prize_Process.bpmn
- Order_Process_for_Pizza_V4.bpmn
- Travel_Booking.bpmn

```{r}
# File from the package
bpmn_file <- system.file("examples/Order_Process_for_Pizza_V4.bpmn", package = "bpmnVisualization")
```


```{r}
# File from the package
#bpmn_file <- system.file("examples/Nobel_Prize_Process.bpmn", package = "bpmnVisualization")
```


```{r}
# File from the package
#bpmn_file <- system.file("examples/Travel_Booking.bpmn", package = "bpmnVisualization")
```


```{r}
# File from the package
#bpmn_file <- system.file("examples/Email_Voting.bpmn", package = "bpmnVisualization")
```

Or
```{r}
#bpmn_file <- file.choose()
```

Or
```{r}
bpmn_file <- here::here("data-raw", "CBI.bpmn")
```

Or
```{r}
#bpmn_file <- "<?xml version="1.0" encoding="UTF-8"?> ...xml_content"
```




### Display the BPMN diagram
```{r}
bpmnVisualization::display(bpmn_file)
```


### Display the BPMN diagram with overlays

```{r}
# overlays <- list(bpmnVisualization::create_overlay("bpmn_element_id_1", "42"), 
#                  bpmnVisualization::create_overlay("bpmn_element_id_2", "9"))
# bpmnVisualization::display(bpmn_file, overlays)
```


### Integrate in Shiny Applications

The following displays a BPMN diagram provided as an example by the package with an overlay on top of a BPMN element.

```{r}
# Install and load the shiny package
# install.packages("shiny")
# library(shiny)
# 
# 
# displayBpmn <- function() {
#     bpmn_file <- system.file("examples/Travel_Booking.bpmn", package = "bpmnVisualization")
#     overlays <- list(bpmnVisualization::create_overlay("_6-203", "9"))
#     bpmnVisualization::display(bpmn_file, overlays)
# }
# 
# ui <- shinyUI(fluidPage(
#     titlePanel("Display bpmn diagrams with execution data"),
#     bpmnVisualization::bpmnVisualizationOutput('bpmnContainer')
#   )
# )
# 
# server = function(input, output) {
#     # renderBpmnVisualization is the R bridge function to the html widgets
#     output$bpmnContainer <- bpmnVisualization::renderBpmnVisualization({ displayBpmn() })
# }
# 
# shinyApp(ui, server)
```

 